Hector Permafrost

### 2019-01-04 analysis/paper/paper.Rmd
Add more notes to paper

Permafrost is an important C reservoir.
There are an estimated 500 Gt C in organic matter buried by frozen loess in the north plains of Siberia and Central Alaska [@zimov_2006_climate_change], and an additional 400 Gt C in other permafrost (CITE; see ref in Zimov), and XXX more in peat (CITE).
This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes (REF).
These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (???) times faster than the global average (REF).
Projections of permafrost C emissions vary.
[@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
Back-of-the-envelope estimates from [@zimov_2006_climate_change]: 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.

Several studies have attempted to evaluate the potential economic impacts of current and future permafrost C emissions.
@hope_2015_economic projected permafrost C emissions using the SiBCASA (TODO: write in full) land surface model and used these projections as inputs to the PAGE09 integrated assessment model (IAM).
They found that XXX.
Similarly, @gonzalez-eguino_2016_significant used the DICE (TODO: write in full) model to XXX.

@@

@@

# Acknowledgements

>>>>

Permafrost is an important C reservoir.
Total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes and wetlands [@turetsky_2002_boreal; @wickland_2006_effects].
But emissions from wet soils may be offset by high organic matter accumulation rates [@camill_2001_changes].
Permafrost thaw in boreal peatlands in north-central Saskatchewan increased CO2 and CH4 fluxes from soil to atmosphere by 1.6 and 30 times, respectively [@turetsky_2002_boreal].
These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (REF).
Projections of permafrost C emissions vary.
[@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
Back-of-the-envelope estimates from [@zimov_2006_climate_change]: 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.

Several studies have attempted to evaluate the potential economic impacts of current and future permafrost C emissions.
@lawrence_2012_simulation project declines in near-surface permafrost by 33% and 72% in RCP 2.6 and 8.5, respectively, and permafrost C feedback will stabilize at 2100 in RCP 2.6.
More recently, @hope_2015_economic projected permafrost C emissions using the SiBCASA (TODO: write in full) land surface model and used these projections as inputs to the PAGE09 integrated assessment model (IAM).
They found that XXX.
Similarly, @gonzalez-eguino_2016_significant used the DICE (TODO: write in full) model to XXX.

@@

Rate of permafrost C release also depends on soil moisture conditions -- drier soils release C much faster ("carbon bomb") than wetter soils ("carbon fizz") [@elberling_2013_long].
Moisture will also affect the balance of aerobic (CO2 release) vs. anaerobic (CH4) C release [@turetsky_2002_boreal], to the extent that unclear if anaerobic (wet) areas are C sources or sinks [@wickland_2006_effects].
Effects of permafrost thaw on soil moisture are a complex hydrological problem -- drainage very sensitive to local (micro-)topography [@wickland_2006_effects].
So will vegetation cover [@wickland_2006_effects].

@@

# Acknowledgments

Funded by EPA grant XXX.
Cyberinfrastructure support from Pacific Northwest National Laboratory (PNNL).
